\chapter{Methodology} \label{chap:method}


In this section, we explain the methoddology used in this thesis. 

Add a flow chart that summarizes the different steps in the pipeline, like the one in the interview presentation.

Add a table showing the cosmological parameters, theri fiducial value and their prior. 




\section{Modelling the electron power spectrum} \label{meth_sec:pee}

explain things in a schematic representation first. 

- a plot for the comparison between Limber and non-Limber approximations.
- a plot validating the number of chosen z and k grid for the power spectrum.
- talk about eh power spcetrum extrapolation.


\subsection{The baryonification correction model} \label{meth_sec:bfc}
explain how we ignore the redshift evolution of some parameters.









\begin{table}[h]
\centering
\begin{tabular}{c c p{6cm} c}
\hline\hline
\textbf{Parameter} & \textbf{Prior} & \textbf{Description} & \textbf{Fixed value} \\
\hline
$\theta_{\mathrm{co}}$ & U[0.001, 0.5] & Core radius of the gas profile relative to $r_{200}$ (Eq. 2.6) & 0.3 \\
$\log_{10} M_{\mathrm{c}}$ & U[11.0, 15.0] & Pivot mass for the mass dependent slope of the gas profile (Eq. 2.6 \& Eq. 2.7) & - \\
$\mu$ & U[0.0, 2.0] & Transition steepness of mass dependent gas parameter (Eq. 2.6 \& Eq. 2.7) & - \\
$\delta$ & U[1.0, 11.0] & Characterizes the transition from inner to outer part of the gas profile (Eq. 2.6) & - \\
$\eta$ & U[0.0, 0.5] & Total stellar fraction within a halo (Eq. 2.10) & 0.1 \\
$d\eta$ & U[0.0, 0.5] & Fraction of central galaxies w.r.t. total stellar fraction (Eq. 2.10) & 0.22 \\
$N_{\mathrm{star}}$ & U[0.0, 0.05] & Stellar fraction amplitude (Eq. 2.10) & 0.030 \\
$c_{\mathrm{iga}}$ & U[0.0, 1.0] & Cold gas fraction (see Eq. 2.12) & 0.1 \\
$b_{\mathrm{hse}}$ & $\mathcal{N}(0.26, 0.07)$ & Hydrostatic mass bias & - \\
$\alpha_{0,\mathrm{nt}}$ & - & Non-thermal pressure amplitude (see Eq. 2.15) & 0.1 \\
\hline\hline
\end{tabular}
\end{table}




\subsubsection{The Core Cosmology Library (CCL)}
describe CLASS and what it does.


\section{Simulating the dispersion measure signal} \label{meth_sec:sim_dm}
Add a table with the fiducial values and their prior. explain how we ignore the redshift evolution of some parameters. Motivate the choice for the number of shells and the redshift range.

\subsection{The Generator for Large Scale Structure (GLASS)}
mention the reason behind the choice of 30 shells. and the binning in comoving space.

\subsection{The electron angular power spectrum} \label{meth_sec:angular_power_spectrum}
\subsection{The numerical implementation of the cosmological covariance matrix}
- Add a plot for the validity of the divisions when implementing in z.



\section{Simulating the systematic uncertainties}

\subsection{Milkyway contribution}
\subsection{Host contribution}

\subsection{Redshift uncertainty}
I simulate the imapct of redshift uncertainty by assuming that a redshift measuremnet $z_i$ is sampled from a Gaussian distribution with mean $z_i$ and standard deviation $\sigma_i = \eta z_i$, where $\eta_i$ is a number between $0$ and $1$. To propagate the error on the cosmological covariance, consider that $(z_i, z_j)  \sim \mathcal{N}((\hat{z}_i, \hat{z}_j) , \Sigma)$, with $\mathcal{N}$ is the normal distribution and
\begin{equation}
    \Sigma = \begin{bmatrix}
    \sigma_i^2 & 0 \\
    0 & \sigma_j^2 
    \end{bmatrix}
\end{equation}
To propagate the error, I calcualte the expectation value $\langle \hat{C}_{ij} \rangle$, given as
\begin{equation}
    \langle \hat{C}_{ij} \rangle = \int \mathcal{N}((\hat{z}_i, \hat{z}_j) , \Sigma) \, \, C_{ij} \, \,dz_i \, \, dz_j
\end{equation}
With Taylor's expansion around $(\hat{z}_i, \hat{z}_j)$ and setting $\Delta_{ij} = (z_i, z_j) - (\hat{z}_i, \hat{z}_j)$, the covariance matrix can be written
\begin{equation}
    \langle \hat{C}_{ij} \rangle = \int \mathcal{N} \, \, \left(\hat{C}_{ij} + \nabla C_{ij}|_{(\hat{z_i}, \hat{z_j})} \Delta_{ij} + \frac{1}{2} \Delta_{ij}^{T} \nabla^2 C_{ij}|_{(\hat{z_i}, \hat{z_j})}\Delta_{ij} + .....\right)\, \,dz_i \, \, dz_j.
\end{equation}
Since the Gaussian distribution has vanaishing moments for moment larger than third order, the expression is simplified into
\begin{equation}
    \langle \hat{C}_{ij} \rangle = \hat{C}_{ij} + \frac{1}{2} \, \, \sigma_{i}^2 \, \, \partial^2_{ii} \, C_{ij}|_{(\hat{z_i}, \hat{z_j})} \, \, + \frac{1}{2} \, \, \sigma_{j}^2 \, \, \partial^2_{jj} \, C_{ij}|_{(\hat{z_i}, \hat{z_j})}.
\end{equation}
Therefore, it is possible to account for the error in the covariance due to redshift uncertainty using the second derivative of the covariance with respect to redshift $z$.

I show the impact of redshift uncertainty on the scatter in the DM-$z$ relation in chapter \ref{chap:results}.


\subsection{Selection effects}
- show a plot of full PDF and a truncated PDF.

\subsection{Unlocalised hosts}

\section{Data Compression}
I simulate the electron density shells and place FRBs in the same redshift and angular position as in the mock catalog. This will reduce the computational cost and the need to calculate the covariance matrix for each realisation within the SBI pipeline. (Try to do this excercise without choosing the sdame redshifts for a lower number of sources such that it lowers down the computational costs)
\subsubsection{Numerical derivative stability}
\subsubsection{The Sensitivity of compressed statistics to model parameters}

Add a plot about derivative stability. Add a plot for the sensitivity.

\section{Simulation-based inference} \label{meth_sec:sbi}
\subsection{Approximate Baysian inference}
\subsection{Neural Posterior Estimator}
\subsection{The Sequentual Neural Posterio Estimator}
\subsubsection{The truncated sequential neural posterior estimator}
\subsection{Flow matching posterior estimator}









