% !TEX root = sbi.tex

%==============================================================================
\chapter{Introduction}
\label{sec:intro}
%==============================================================================
The questions on how the universe came to be, why the universe is as it is, and what the universe is heading towards fall upon cosmology to answer. Cosmology is the study of the universe on large scales in an attempt to answer the latter questions {\color{green} cite maybe Peebles}. This pursuit started as data-starved research {\color{green} cite} and transformed into a data-rich endeavour as a result of a plethora of imaging surveys such as the LSST Vera Rubin observatory, Euclid, and Nancy Roman {\color{green} cite}. The qualitative questions, with which this paragraph started, translate in the statistical sense to be what is the cosmological model that best describes the universe across all length scales and energy scales.

The first attempt to describe the universe as a whole came through the lens of the general theory of relativity where Robertson and Walker  {\color{green} cite} built on the work of Friedmann and Lemaître to describe the geometry of the universe through what is now known as the Friedmann-Lemaître-Robertson-Walker (FLRW) metric. We discuss this metric further in \ref{theory_sec:FLRW}. This metric was largely motivated by the observations of Edwin Hubble who saw that galaxies further away from the Earth are moving faster than galaxies closer to Earth \cite{Hubble_1929}. Figure shows the plot from Hubble's work in 1929.

Further observations of the magnitude of distant galaxies have shown that the universe is expanding {\color{green} cite}. The general theory of relativity attributes this expansion to the energy density content of the universe. Therefore, the question of what is driving the expansion of the universe and what are the components that contribute to the energy budget of the universe represent a crucial question in cosmology for which multiple cosmological models have been suggested \citep{di2025cosmoverse}. In fact, the cosmological model has been motivated by both cosmological and astrophysical observations. Some of these observations are the measurement of galaxy rotation curves by Vera Rubin {\color{green} cite} that hinted at an additional matter component that does not interact with light and the discovery of the cosmic microwave background (CMB) radiation that hinted at a hot and homogeneous early universe.{\color{green} cite}. Furthermore, early observations of clusters of galaxies have shown that the spatial distribution of galaxies deviates from randomness. This was shown by Peebles {\color{green} cite}. By calculating the two point correlation function, there was clearly a signal in the way that galaxies are distributed which points out to an underlying mechanism in the formation of these galaxies \citep{Totsuji_Kihara_1969}. The development of observational techniques and instrumentation allowed the launch of the Center for Astrophysics Redshift Survey \citep{Davis_Huchra_Latham_Tonry_1982} which showed that galaxy positions are correlated on large scales and that there is a structure in the universe \citep{Davis_Peebles_1983}.

In the early 1990s, cosmology experienced a paradigm shift with the launch of the Cosmic Background Explorer (COBE) satellite \citep{Boggess_Mather_etal._1992}. The COBE mission provided the first detection of primordial CMB anisotropies and further confirmed that the universe is flat \citep{Smoot_1999}. In the context of the FLRW metric, the flatness of the universe requires a total sum of the relative density parameters to be 1 (within error bounds). Therefore, in addition to the matter component (dark and baryons), there should be a third component that can account for the missing energy density. Using supernova data from the Hubble Space Telescope \citep{Bahcall_Spitzer_1982}, \citet{Riess_etal._1998} found strong evidence for an expanding universe with negative acceleration and supported the findings of the COBE mission for the existence of an additional density component different from matter which is driving the expansion of the universe. Hereafter, the term `dark energy' was coined to refer to this component \citep{Turner_2001}. With this, the $\Lambda$CDM model was born, and the efforts to obtain precise estimates of the cosmological parameters intensified \citep{Bennett_etal._2003, Pope_2004}.

While early experiments relied on CMB, galaxy number counts, or supernova to infer cosmological parameters, gravitational lensing was in development as a tool to constrain the cosmological parameters. Gravitational lensing is the distortion of light geodesics by gravity when light travels from a distant source through the inhomogeneous universe towards an observer on the Earth \citep{Schneider_Ehlers_Falco}. On the level of a galaxy, gravitational lensing effects manifest themselves as a distortion of the shape of the galaxy and a change in the angular position with respect to a particular observer. \citet{Kaiser_Squires_1993} showed that these distortions could be used to obtain a map of the integrated mass density along the line of sight. The use of galaxy shape distortions for cosmology was a direct result of the Kaiser-Squires inversion formula where \citet{seitz1996steps} realised that the expectation value of galaxy shape is an unbiased estimate of the shear in the weak field limit. In principle, it is possible to relate the power spectrum of galaxy shapes with the matter power spectrum, offering a concrete method of probing cosmology by studying the correlations between galaxy shapes \citep{Bartelmann_Schneider_2001}. This new probe of the large scale structure is called cosmic shear and was used by \citet{Maoli_VanWaerbeke_2001} to obtain constraints on the cosmological parameters. These constraints were not as tight as the ones obtained by COBE, however, it paved the road towards using cosmic shear to constrain the cosmological model. The advantage of cosmic shear as an integrated quantity along the line of sight means that it is sensitive to structure growth in the late universe. Therefore, it represents a very good tool to constrain the equation of state for dark energy \citep{Hu_Tegmark_1999, Jain_Taylor_2003}.

After that, multiple galaxy surveys used cosmic shear for cosmological parameter inference, and numerous statistical tools were developed to deal with the increasing data size \citep{Schneider_VanWaerbeke_Jain_Kruse_1998, Schneider_Eifler_Krause_2010}. Likewise, with increasing data size, the noise level in a cosmological signal increases and the need to mitigate the impact of the systematic and measurement uncertainties is imperative. In this work, we focus on tackling one of the most crucial systematics, namely baryonic feedback on the large scale structure.

Baryonic feedback are the set of baryonic physics processes that impact the distribution of matter \citep{Rudd_Zentner_Kravtsov_2008}. These processes include, but are not limited to, supernova feedback, active galactic nuclei feedback, and star formation. By analysing hydrodynamical simulations,  \citet{van2011effects} realised that baryonic feedback affects the matter distribution in the universe at different scales, meaning that their impact should be modelled and included in a theory of structure formation. However, the theoretical formalism of structure growth, and therefore the matter power spectrum, assumes that structure growth is only driven by gravitational effects, ignoring baryonic feedback \citep{Dodelson_Schmidt_2026}. Modelling baryonic feedback analytically is a cumbersome task, especially since the physics of galaxy formation and evolution is highly uncertain where different suites of hydrodynamical simulations have different feedback models, and each one of these models yields a totally different description of the matter power spectrum suppression \citep{van2011effects}. The mismodelling of the matter power spectrum tunnels down to the modelled cosmic shear signal and yields biased parameter estimates in a cosmological analysis.

On a conquest for precision cosmology, multiple statistical and modelling methods attempted to mitigate the impact of baryonic feedback on the large scale structure. For example, \citet{krause2021dark} imposed scale cuts on the measured shear correlation function and only considered larger scales where baryonic feedback is minimal. However, the cosmic shear signal derives most of its sensitivity from non-linear scales, i.e. small scales \citep{Piccirilli_Zennaro_García-García_Alonso_2025}. Therefore, imposing a scale cut leads to loss of information. Other statistical techniques involve the use of principal component analysis to compress the parameter space of baryonic feedback parameters from hydrodynamical simulations and marginalising over these parameters in a cosmological analysis \citep{Eifler_Krause_Dodelson_Zentner_Hearin_Gnedin_2015}. Another statistical method involves introducing an additive term to the covariance matrix called the theoretical error covariance \citep{Moreira_2021, Maraio_2025a} which describes the variation of different baryonic feedback scenarios among all available hydrodynamical simulations. Analyses involving all the aforementioned results yield unbiased estimates of the cosmological parameters, however, it comes at a cost of introducing additional scatter due to the existing nuisance parameters. Finally, \citet{Schneider_Teyssier_2015} introduced the baryonification correction model, a model that parametrizes the displacement of dark matter particles in the existence of baryonic feedback. This model and its developments \citep{Schneider_Kovač_etal._2025} were calibrated on hydrodynamical simulations and were able to emulate baryonic feedback in other hydrodynamical simulations that they were not calibrated on. Having this model, it is possible to obtain a prediction for any probe of baryons in the universe. In particular, this model was used to constrain baryonic feedback parameters with X-ray and the kinematic Sunyaev-Zel'dovich effect \citep{Kovač_2025} since it is possible to obtain a prediction of gas power spectra from which a simulated signal of the corresponding baryon tracer could be obtained. Though these probes are promising, they rely on some assumptions with respect to the structure of the galaxy clusters such as having hydrodynamic equilibrium and require many convolutions from a prediction of the gas power spectrum up to a predicted signal. Luckily, fast radio bursts provide a direct probe of the electron density in the universe through the information that their signal encodes.

Fast radio bursts (FRBs) are pulses of electromagnetic waves of unknown extragalactic origin. First discovered in 2007 \citep{Lorimer_Bailes_McLaughlin_Narkevic_Crawford_2007}, the most probable canditate for the sources of FRBs is the magnetar which is a star remenant that underwent a supernova. The signal of FRBs gets dispersed while travelling through the universe, meaning that the wave packets of lower frequency will arrive later than wave packets of higher frequency. This difference in arrival time is directly proportional to the integrated electron density along the line of sight where the dispersion measure signal could be easily simulated with minimal assumptions about the relationship of the dispersion measure and the electron number density in the universe. The grand benefit of the dispersion measure signal lies in its constraining power of baryonic feedback parameters since it is a direct tracer of the baryon distribution in the universe. Even though the source of these bursts is unknown \citep{Petroff_2019}, it is still possible to exploit the dispersion measure signal in a cosmological context, given that FRBs are hosted within galaxies whose redshift can be identified.

Simulated likelihood analyses showed that the cross-correlation of FRBs with other probes of structure yields better constraints on some cosmological parameters. In particular, calibrating baryonic feedback with FRBs and weak lensing led to one order of magnitude better constraints on the baryonic feedback and 50\% improvements on constraints of the sum of neutrino masses $ \sum m_\nu$ and the baryonic fraction in the universe $\Omega_b$ \citep{reischke2023calibrating}. Furthermore, combining FRBs with galaxy number counts in a 3x2-point analysis provided better constraints on baryonic feedback parameters as compared to galaxy number counts \citep{Sharma_2026}. In practice, we can aim to use FRBs alone as a probe to impose tight constraints on the baryonic feedback parameters and use these constraints as priors in a cosmological analysis involving probes of structure, allowing the information content of LSS data to constrain the cosmological parameters and delegating the task of constraining baryonic feedback to FRBs.

The promising potential of FRBs in constraining baryonic feedback could be hindered by multiple modelling and measurement uncertainties. First, FRBs are observed using radio telescopes with only limited follow-up optical observations that are necessary to calculate the redshifts of galaxies and in turn quantify the associated mean DM at that redshift. Second, measurements of these redshifts are highly uncertain especially for sources at high redshifts. Third, the measured dispersion measure signal receives three main contributions: a contribution from the Milky Way, a contribution from the host galaxy, and a contribution from the circumgalactic medium which is the quantity of cosmological interest \citep{Reischke:2025srr}. The models for both the Milky Way and the host galaxy contribution are both uncertain, leading to biases in the parameter constraints if not dealt with correctly. Fourth, the theoretical formalism of fast radio bursts assumes that the radio signal travels in a homogeneous universe, ignoring any gravitational lensing effect due to large scale structure. There are contradicting results on the impact of gravitational lensing; \citet{alonso2021linear} showed that first order contributions from general relativistic effects to the DM angular power spectrum are subdominant compared to the contribution from the electron column density, meanwhile, \citet{Takahashi_2024} showed that the second order contribution due to lensing could lead to a 1\% deviation in the luminosity distance, resulting in an underestimation of the luminosity distance.

Therefore, the purpose of this thesis is to model, forecast, and explore mitigation strategies of systematic uncertainties and describe their importance with varying sample size. With upcoming radio surveys, a plethora of data will be available such that an accurate modelling of the dispersion measure signal is crucial to extract meaningful information from the signal \citep{LukeHoffmann_James_Glowacki_Prochaska_Gordon_Deller_Shannon_Ryder_2024}. I scale up previous work by \citet{Konar_Reischke_Hagstotz_Nicola_Hildebrandt_2025a} by using simulation-based inference (SBI) to infer the impact of systematic uncertainties on the baryonic parameter constraints using fast radio bursts only. Using SBI in this context has multiple benefits. First, it facilitates integrating different uncertainty effects on the level of the dispersion measure signal either on the level of the dispersion measure or the level of the redshift. Second, it removes the necessity to approximate the likelihood of the data given the parameters, unlike traditional Bayesian analyses. Results from ray tracing in hydrodynamical simulations \citep{Konietzka:2025kdr} suggest that a Gaussian or lognormal approximation to the probability distribution function breaks down for large numbers of FRBs, and an analytic model of the dispersion measure is needed to simulate the PDF of FRBs \citep{torkamani2026baryonification}. More details on SBI and the method used in this thesis can be found in \ref{meth_sec:sbi}

This thesis is structured as follows: I review the theoretical background in \ref{chap:theory} including the general theory of relativity (\ref{theory_sec:GR}), the FLRW metric (\ref{theory_sec:FLRW}), the cosmological model (\ref{theory_sec:cosmo_model}), weak gravitational lensing (\ref{theory_sec:WL}), baryonic feedback on the large scale structure (\ref{theory_sec:BF}), the basics of fast radio bursts (\ref{theory_sec:FRBs}), and the systematics uncertainties in modelling the DM signal (\ref{theory_sec:uncertainties}). I describe the methodology in \ref{chap:method} including modelling the electron power spectrum (\ref{meth_sec:pee}), simulating the dispersion measure signal (\ref{meth_sec:sim_dm}), describing the systematic uncertainties (\ref{meth_sec:uncertainties}), and explaining simulation-based inference (\ref{meth_sec:sbi}). In \ref{chap:results}, I describe the results. Finally, I discuss the results in \ref{chap:discuss}
