% !TEX root = sbi.tex

%==============================================================================
\chapter*{Acknowledgements}%
\label{sec:ack}
%==============================================================================

I am strongly indebted to my supervisors Dr. Andrina Nicola and Dr. Robert Reischke for their immense support throughout the duration my master's studies. Their support and advice were pivotal in the process of conducting this work. I would also like to thank the member of the 'Combining Cosmological Probes' group at the Argelander Institute for Astronomy, especially, Subarna, Mokka, Micheal, Reihana, and Reiko.

I would like to thank Prof. Dr. Jennifer Schöber for stepping in to be my supervisor for the second part of my thesis.

Furthermore, I would like to thank my friends and colleagues at the University of Bonn who made the master's studies journey more fun and less stressful. I would like to thank my aunt, Sana, who kept my company and offered immense support when adjusting to the life in Germany and Bonn. 

I thank my girlfriend, Fée, who was my best friend throughout the journey of my master's thesis. Her words of encourgments were crucial in the moments of doubt. 

I would further like to thank the memebers of the physical cosmology department at the Max Planck Institute for Astrophysics where part of this work was conducted, in particular, Fabian Schmidt, Šafak Celik, Lena Scheuchel, Sveva Castillo, Florian List, and Adr

You should probably use \texttt{\textbackslash chapter*} for
acknowledgements at the beginning of a thesis and
\texttt{\textbackslash chapter} for the end.
